Ligeti, Stockhausen and Koenig:\\Aesthetics of Electronic Music

### 2015-09-26 manuscript.tex
Ramifications analysis

\subsection{Comparison}

@@

The last concept is the representation of pitch ranges.
It is clearly a serial concept, but part of it, especially the ``transitional'' representation, remind me the ``Mistiness'' to ``clearing-up'' process of Ligeti.
Extract this idea from its serial way of thinking, and it can be seen as a very expressive description of relationships between ``planes'' of sounds and melodies, including intermediate states.

\section{Ligeti's Remifications}
\label{sub:ramifications}

Ligeti ideals and characteristics as found in Remifications.

@@

This ideas are perhaps the most dominants in this composition.
Examples can be found in the 0:00 - 1:55, 4:20 - 4:35, 5:28 - 6:25.

@@

It appears to me that these idea could be counted together with the previous ones.
On the other hand, some parts in Remifications are dominated by more ``machine-like'' characteristics than others.
For example: 5:28 - 6:25, 7:00 - 8:00.

Gesticulated phrases appears in Remifications only for short periods, they are not characterize the composition as a hole.
Nevertheless, we can note two ``overly exaggerated'' phrases, where the primer is a bit more distinct in that manner: 3:36 - 3:45, 6:28 - 6:40.
One could argue that these are just dramatic phrases within the composition, but I suspect that this amount of vibrato in Ligeti's composition most present some decent amount of cynicism.
TODO cite frigyesi to support this statement.

>>>>

\subsection{Comparison}
\label{sub:comparison}

@@

The last concept is the representation of pitch ranges.
It is clearly a serial concept, but part of it, especially the ``transitional'' representation, reminds me the ``Mistiness'' to ``clearing-up'' process of Ligeti.
Extract this idea from its serial way of thinking, and it can be seen as a very expressive description of relationships between ``planes'' of sounds and melodies, including intermediate states.

\section{Analysis of Ligeti's Ramifications}
\label{sub:ramifications}

In this section I analyze Ligeti's Ramifications according to the concepts described before.
The order of the concepts in the analysis will follow the three categories described in the last section (\ref{sub:comparison}), starting from sound, through serializm and finally to higher level expressions.
In each category I analyze general concepts first when applicable.

The analysis was made by hearing a recording of Ramification from 2003 (\cite{rami_music}).
I found the recording more informative than the composition score;
meaning that Ligeti's ideas such as misleading notation and flexible rhythm appears more clearly in the audible result than in the score itself.
Throughout the analysis I refer to timestamps in the piece for examples based on this recording.

\subsection*{Non-stationary sounds of unlimited length}

Ligeti presents these ideas throughout the whole composition.
There are only brief sections where the music does not present an infinite texture of polyphonic lines.
The polyphonic lines, as Ligeti explains, are not meant to be heard by themselves.
Their purpose is to serve the overall texture, the ``woven cobweb''.
Koenig ideas of non-stationary sounds and sounds of unlimited length fits this composition nicely.

\subsection*{Lower level building blocks}

This concept is coupled tightly to the process of electronically creating sounds.
Hence, it is not trivial to extract some fundamental concepts from this process, to be relevant to Ramifications.
Nevertheless, Stockhausen argues that ``one would not proceed from sound properties that had already been experienced'' (\cite[p. 42]{stockhausen}).
In that regard, I think that Ligeti does not relies on already established properties of the sound and their derivatives, but rather explores different low level blocks.
Compared to the historically established melody / harmony building blocks of western music Ligetis' foundations include the already discussed concepts of micropolyphony, flexible rhythm and misleading notation.

@@

Beyond the general concepts above, this ideas are perhaps the most dominants in this composition.
Examples can be found in the 0:00 - 1:55, 4:20 - 4:35, 5:28 - 6:25.

\subsection*{Sound as a formal event}

There are some more consonant sounds in this composition that serve somewhat as a structuring tool;
they divide the composition into sections.
Namely, the first one is found in 2:28, and it marks the end of the first part of the composition and divide it from the rest.
The second obvious sound event is found in 6:45.
It marks the beginning of the last section of the composition.

\subsection*{Noise}

Noise, as Stockhausen describe it, is found in this composition mainly in the rhythm layer, and it might be understood in two different ways:

\begin{itemize}
  \item First, the micropolyphony and flexible rhythm presented in this composition can be understood as a noise like phenomenon;
  the players repetitive melody lines are not part of a clearly structured rhythm;
  the lines might share the same beat for a moment and then gradually shift back to different tempi (1:01 - 1:24 or 5:32 - 6:25 for example).
  \item Second, the absence of audible bars can also be understood as noise.
  The bar lines are a natural part of western music, and take an important role in dividing the melody / harmony into repeatable mid-level structure.
  In Ramification they can't be heard.
  Instead, the mid-level structure is much more dynamic; one cannot anticipate it as with traditional bars.
\end{itemize}

\subsection*{The correlation of timbre, pitch, intensity and duration}

Although this idea can be understood outside of serial thinking, as I argued before, I'm unable to find its occurrence in Ramifications.
This concept is, perhaps, too contradicted to Ligeti's way of thinking.
I will refer to this open issue again later, in the discussion.

\subsection*{Representation of pitch ranges}

This idea shines throughout Ramifications.
The opening of the piece serve as a good example.
From the start, Ligeti `represent' a very narrow pitch range, using a sound plane.
This sound plane is in a transitory mode most of the time;
the sound repeatedly moved back and forth between dense textures and recognizable melodic lines.
In addition, after one minute the plane starts to move and expand its higher bound to higher pitches.
This trend accelerate till 2:19, where, by slowing the patterns and reducing the layer, the plane transitions into a unison.

Gesticulated phrases appears in Ramifications only for short periods, they are not characterize the composition as a whole.
Nevertheless, I can note two ``overly exaggerated'' phrases, where the primer is a bit more distinct for that mater: 3:36 - 3:45, 6:28 - 6:40.
One could argue that these are just dramatic phrases within the composition, but I suspect that this amount of vibrato in Ligeti's composition most represent some decent amount of sarcasm (TODO cite frigyesi).

@@

It appears to me that this idea could be counted together with those presented in the micropolyphony, flexible rhythm and misleading notation above.
On the other hand, some parts in Ramifications are more dominated by ``machine-like'' characteristics than others.
For example: 4:23 - 4:53, 5:28 - 6:25, 7:00 - 8:00.
